\documentclass[
    letterpaper,paper=portrait,fleqn,
    DIV=16,fontsize=12pt,twoside=semi,
    parskip=full-,
    headings=standardclasses]
{scrartcl}

\usepackage{scrlayer-scrpage}
\clearpairofpagestyles
\ohead{\pagemark}

\usepackage{mathtools}
\usepackage[fullfamily,footnotefigures,swash,lf,mathtabular]{MinionPro}
\usepackage[eqno,enum,lineno]{tabfigures}
\usepackage{pifont}

\usepackage{enumitem}
\usepackage[dvipsnames]{xcolor}
\usepackage[capitalise,nameinlink]{cleveref}
\usepackage{booktabs}
\usepackage{ragged2e}
\usepackage{dcolumn}
\usepackage{tikz}

% Reduce vertical spacing around headings
\RedeclareSectionCommand[
  runin=false,afterindent=false,
  beforeskip=0.5\baselineskip,
  afterskip=0.5\baselineskip
]{section}
\RedeclareSectionCommand[
  runin=false,afterindent=false,
  beforeskip=0pt
]{paragraph}
% Allow display-mode math to break across pages
\allowdisplaybreaks[1]
% Set RaggedRight paragraph indent
\setlength{\RaggedRightParindent}{\parindent}
% Create additional column types
\newcolumntype{d}[1]{D{.}{.}{#1}}
\newcolumntype{R}{>{\raggedleft\arraybackslash}p{1in}}
% Bold enumerate numbering
\setlist[enumerate,1]{label=\bfseries\arabic*.}
%\setlist{itemsep=\baselineskip,parsep=0\baselineskip}
% Sections counted in alphabet and formatted as ``Parts''
\renewcommand*\thesection{\Alph{section}}
\renewcommand\sectionformat{Part~\thesection:\enskip}
\crefname{section}{part}{parts}

\begin{document}
\RaggedRight

ECO 120-04 \\
Lucas Reddinger \\
Wednesday 23 November 2022 \hfill Your full name: \underline{\hspace{3.25in}}

\vspace{0.7\baselineskip}
\textbf{\LARGE Assignment 10: Stagflation}
\vspace{0.3\baselineskip}

\emph{Due Monday 28 November.} Please submit hardcopy at the beginning of class (11:00 a.m.), or if you prefer, under the door of Wimberly Hall 339C by 10:50 a.m.

{\centering

\vspace{-0.5\baselineskip}
\Pisymbol{MinionPro-Extra}{121}
\vspace{-0.5\baselineskip}

}

Suppose that the U.S.~economy is initially in long-run equilibrium, producing its potential output $Y_P$, as depicted directly below.

\begin{tikzpicture}[scale=0.5]
\draw[thick,<->] (0,10) node[below left,label={[align=right]left:Aggregate\\price level\\ }] {$P$} --(0,0)--(20,0) node[below left,label=right:Real GDP]{$Y$};
\draw[very thick] (2,8)--(16,1) node[right]{$\text{AD}_1$};
\draw[very thick,orange] (13,0) node[below]{$Y_P$}--(13,8) node[above]{LRAS};
\draw[very thick,red] (11,1)--(19,7) node[right]{$\text{SRAS}_1$};
\draw[dashed] (0,2.5) node[left]{$P_1$}--(13,2.5);
\end{tikzpicture}

A multinational effort to sanction Russian oil and natural gas results in a negative shock to short-run aggregate supply from $\text{SRAS}_1$ to $\text{SRAS}_2$ as depicted below. U.S.~economic output falls from $Y_P$ to $Y_2$, and the aggregate price level rises from $P_1$ to $P_2$. This results in a period of \emph{stagflation}---lower economic output coupled with inflation.

\begin{tikzpicture}[scale=0.5]
\draw[thick,<->] (0,10) node[below left,label={[align=right]left:Aggregate\\price level\\ }] {$P$} --(0,0)--(20,0) node[below left,label=right:Real GDP]{$Y$};
\draw[very thick] (2,8)--(16,1) node[right]{$\text{AD}_1$};
\draw[very thick,orange] (13,0) node[below]{$Y_P$}--(13,8) node[above]{LRAS};
\draw[very thick,red] (11,1)--(19,7) node[right]{$\text{SRAS}_1$};
\draw[dashed] (0,2.5) node[left]{$P_1$}--(13,2.5) ;
\draw[very thick,red] (6,1)--(15,7.75) node[right]{$\text{SRAS}_2$};
\draw[dashed] (0,4) node[left]{$P_2$}--(10,4) --(10,0)node[below]{$Y_2$};
\draw[line width=2pt,->] (18,6.75)--(14.5,6.75);
\end{tikzpicture}

\clearpage

\begin{enumerate}
\item Please illustrate how fiscal policy could be used to maintain economic growth. \\ {\footnotesize Hint: Reproduce the second graph here. Illustrate a change in fiscal policy that restores output to $Y_P$.}

\vfill

\item Please illustrate how fiscal policy could be used to stabilize prices. \\ {\footnotesize Hint: Reproduce the second graph here. Illustrate a change in fiscal policy that restores the price level to $P_1$.}

\vfill

\item Can fiscal policy remedy stagflation? Please explain with a complete sentence.  \\ {\footnotesize Hint: Can a change in fiscal policy restore both output to $Y_P$ and the price level to $P_1$?}

\vspace{2\baselineskip}

\end{enumerate}

\end{document}