\documentclass[
    letterpaper,paper=portrait,fleqn,
    DIV=16,fontsize=12pt,twoside=semi,
    parskip=full-,
    headings=standardclasses]
{scrartcl}

\usepackage{scrlayer-scrpage}
\clearpairofpagestyles
\ohead{\pagemark}

\usepackage{mathtools}
\usepackage[fullfamily,footnotefigures,swash,lf,mathtabular]{MinionPro}
\usepackage[eqno,enum,lineno]{tabfigures}

\usepackage{enumitem}
\usepackage[dvipsnames]{xcolor}
\usepackage[capitalise,nameinlink]{cleveref}
\usepackage{booktabs}
\usepackage{ragged2e}
\usepackage{dcolumn}
\usepackage{tikz}

% Reduce vertical spacing around headings
\RedeclareSectionCommand[
  runin=false,afterindent=false,
  beforeskip=0.5\baselineskip,
  afterskip=0.5\baselineskip
]{section}
\RedeclareSectionCommand[
  runin=false,afterindent=false,
  beforeskip=0pt
]{paragraph}
% Allow display-mode math to break across pages
\allowdisplaybreaks[1]
% Set RaggedRight paragraph indent
\setlength{\RaggedRightParindent}{\parindent}
% Create additional column types
\newcolumntype{d}[1]{D{.}{.}{#1}}
\newcolumntype{R}{>{\raggedleft\arraybackslash}p{1in}}
% Bold enumerate numbering
\setlist[enumerate,1]{label=\bfseries\arabic*.}
%\setlist{itemsep=\baselineskip,parsep=0\baselineskip}
% Sections counted in alphabet and formatted as ``Parts''
\renewcommand*\thesection{\Alph{section}}
\renewcommand\sectionformat{Part~\thesection:\enskip}
\crefname{section}{part}{parts}

\begin{document}
\RaggedRight
\thispagestyle{plain}

ECO 120-04 \\
Lucas Reddinger \\
Wednesday 21 November 2022 \hfill Your full name: \underline{\hspace{3.25in}}

\vspace{0.7\baselineskip}
\textbf{\LARGE Assignment 9: Aggregate supply and demand}
\vspace{0.3\baselineskip}

\emph{Due Monday 28 November.} Please submit hardcopy at the beginning of class (11:00 a.m.), or if you prefer, under the door of Wimberly Hall 339C by 10:50 a.m.

\section{Recessions and expansions}

\begin{enumerate}[font={\bfseries}]
\item Please define a recessionary period using actual output $Y$ and potential output $Y_p$.
\vfill
\item Please define an expansionary period using actual output $Y$ and potential output $Y_p$.
\vfill
\item Please define the output gap using actual output $Y$ and potential output $Y_p$.
\vfill
\item Please define a recessionary period using the output gap.
\vfill
\item Please define an expansionary period using the output gap.
\vfill
\end{enumerate}

\vspace{-2\baselineskip}
\clearpage

\section{An economy during a recessionary period\label{sec:recession}}

Please graph an economy in a recessionary period using the aggregate supply and demand model.

\begin{enumerate}[font={\bfseries}]
\item Begin by drawing and labeling axes.
\item Draw and label an aggregate demand (AD) and short-run aggregate supply (SRAS) curves.
\item Label the aggregate price level $P_0$ and the actual output level $Y_0$.
\item In a recession, is potential output $Y_p$ above or below actual output $Y_0$? \hfill \underline{\hspace{1.5in}}
\item Add a long-run aggregate supply (LRAS) curve to your graph, and label $Y_p$.
\end{enumerate}

\vfill

\begin{center}
\begin{tikzpicture}[x=3mm, y=3mm]
\clip(3,3) rectangle (52,52);
\draw[step=4mm, very thin, black!20!white] (0,0) grid (60,120);
\draw[step=20mm, thick, black!40!white] (0,0) grid (60,120);
\end{tikzpicture}
\end{center}

\clearpage

\section{An economy during an expansionary period\label{sec:expansion}}

Please graph an economy in an expansionary period using the aggregate supply and demand model.

\begin{enumerate}[font={\bfseries}]
\item Begin by drawing and labeling axes.
\item Draw and label an aggregate demand (AD) and short-run aggregate supply (SRAS) curves.
\item Label the aggregate price level $P_0$ and the actual output level $Y_0$.
\item In an expansion, is potential output $Y_p$ above or below actual output $Y_0$? \hfill \underline{\hspace{1.5in}}
\item Add a long-run aggregate supply (LRAS) curve to your graph, and label $Y_p$.
\end{enumerate}

\vfill

\begin{center}
\begin{tikzpicture}[x=3mm, y=3mm]
\clip(3,3) rectangle (52,52);
\draw[step=4mm, very thin, black!20!white] (0,0) grid (60,120);
\draw[step=20mm, thick, black!40!white] (0,0) grid (60,120);
\end{tikzpicture}
\end{center}

\section{The short-run and the long-run}

\begin{enumerate}[font={\bfseries}]

\item What characterizes the short-run in the aggregate supply and demand (AS and AD) model?

\vfill

\item Which curve shifts as nominal wages change?

\vfill

\item In which direction does the curve shift \emph{when nominal wages increase}?

\vspace{1.0\baselineskip}
The \underline{\hspace{2in}} curve shifts \underline{\hspace{2in}}.
\vspace{1.0\baselineskip}

\item In which direction does the curve shift \emph{when nominal wages decrease}?

\vspace{1.0\baselineskip}
The \underline{\hspace{2in}} curve shifts \underline{\hspace{2in}}.
\vspace{1.0\baselineskip}

\item In long-run equilibrium, $Y=Y_p$. Revisit \cref{sec:recession}, with \emph{an economy in recession}. Suppose that wages adjust so that the economy returns to long-run equilibrium. Consider what shift would be necessary to make $Y=Y_p$. Graph this shift, and label the new price level $P_1$ and new actual output level $Y_1$. Did wages increase or decrease?

\vfill

\item In long-run equilibrium, $Y=Y_p$. Revisit \cref{sec:expansion}, with \emph{an economy in expansion}. Suppose that wages adjust so that the economy returns to long-run equilibrium. Consider what shift would be necessary to make $Y=Y_p$. Graph this shift, and label the new price level $P_1$ and new actual output level $Y_1$. Did wages increase or decrease?

\vfill

\vspace{-2\baselineskip}
\clearpage

\end{enumerate}

\end{document}
